---
title: "R Notebook"
output:
  pdf_document: default
---


```{r}
library(sf)
library(stars)
library(ggspatial)
library(scales)
library(grid)
library(forcats)
library(tidyverse)
```

## Load data

```{r}
# read the NLCD data, pull out its coordinate system
# note this is a stars_proxy object because it's too large
nlcd <- stars::read_stars('/mnt/data/nlcd2016_ny.tif')
nlcd_coords <- st_crs(nlcd)

nlcd_types <- read_csv('/mnt/rmd/NLCD_classes.csv') %>% 
  mutate(description = fct_reorder(description, code),
         code = as.factor(code))

# substations
substations <- sf::read_sf('/mnt/data/usa_electric_substations.kml')
substations_ny <- filter(substations, STATE == "NY", STATUS == "IN SERVICE") %>% 
  st_transform(nlcd_coords)

# state boundaries
states <- sf::read_sf('/mnt/data/usa_states.geojson.json')
ny_boundary <- filter(states, NAME == "New York") %>% 
  st_transform(nlcd_coords)

# pull out ny state boundaries untransformed for use next
ny_boundary_latlng <- filter(states, NAME == "New York")

# transmission lines
transmission_lines_ny <- sf::read_sf('/mnt/data/Electric_Power_Transmission_Lines.geojson') %>%
  filter(VOLTAGE >= 230) %>% #230 kV on up means high voltage transmission; note some are missing voltage, they're excluded
  st_intersection(ny_boundary_latlng) %>%
  st_transform(nlcd_coords)
```

Basic map of substations within NY state
```{r}
ggplot() + layer_spatial(ny_boundary) + layer_spatial(substations_ny) + layer_spatial(transmission_lines_ny)
```

## Buffer

Make a two mile buffer around all substations, 1 mile around transmission lines
```{r}
substation_buffer <- st_buffer(substations_ny, units::as_units('2mi'))
transmission_line_buffer <- st_buffer(transmission_lines_ny, units::as_units('1mi'))
buffer_together <- bind_rows(list(transmission_line_buffer %>% select(geometry), 
                                  substation_buffer %>% select(geometry))) %>%
                   st_union()
```

What percentage of NY state land is covered by our buffer zone?

```{r}
st_area(buffer_together) / st_area(ny_boundary)
```


## Land use types in buffer

Pull out colors for future use
```{r}
nlcd_colors <- pull(nlcd_types, hex_color)
names(nlcd_colors) <- pull(nlcd_types, description)
```

Zoom in for a more local look
```{r}
ossining = c(xmin=-73.97, ymin=41.09, xmax=-73.69, ymax=41.33)
bbox <- st_bbox(ossining, crs=4326) %>% 
      st_as_sfc() %>% 
      st_transform(nlcd_coords)
nlcd_zoom <- st_crop(nlcd, bbox) %>% st_as_stars() # crop and change from proxy to full stars
```

```{r}
nlcd_zoom %>%
  as_tibble() %>%
  mutate(code = as.factor(nlcd2016_ny.tif)) %>%
  left_join(nlcd_types, by = 'code') %>%
  ggplot(aes(x=x, y=y, fill=description)) + 
  geom_raster() + 
  scale_fill_manual(values=nlcd_colors) +
  layer_spatial(transmission_lines_ny %>% st_crop(bbox)) + 
  layer_spatial(substations_ny %>% st_crop(bbox))
  
```

```{r}
nlcd_zoom[buffer_together] %>%
  as_tibble() %>%
  mutate(code = as.factor(nlcd2016_ny.tif)) %>%
  left_join(nlcd_types, by = 'code') %>%
  ggplot(aes(x=x, y=y, fill=description)) + 
  geom_raster() + 
  scale_fill_manual(values=nlcd_colors) +
  layer_spatial(transmission_lines_ny %>% st_crop(bbox)) + 
  layer_spatial(substations_ny %>% st_crop(bbox))
```
Zoom out to all of NY State

```{r}
nlcd[buffer_together] %>%
  st_as_stars(downsample=10) %>%
  as_tibble() %>%
  mutate(code = as.factor(nlcd2016_ny.tif)) %>%
  left_join(nlcd_types, by = 'code') %>%
  ggplot(aes(x=x, y=y, fill=description)) + 
  geom_raster() + 
  scale_fill_manual(values=nlcd_colors)
```


```{r}
# via https://github.com/grattan/grattantheme/blob/master/R/watermark.R
watermark <- function(watermark, fontsize = 120,
                          colour = "grey90", alpha = 0.05,
                          fontface = "bold", angle = 22) {
  watermark_grob <- textGrob(watermark, gp = gpar(fontsize = fontsize, colour = colour,
                                      alpha = alpha, fontface = fontface),
                   rot = angle)

  annotation_custom(grob = watermark_grob)
}
```

To speed things up, we work with a downsampled grid, doesn't make any difference for the relative proportions at this scale
```{r}
nlcd_downsample <- 10
```

```{r}
nlcd[buffer_together] %>%
  pull(nlcd2016_ny.tif) %>%
  st_as_stars(downsample = nlcd_downsample) %>%
  table() %>%
  as_tibble() %>%
  rename(code = 1) %>%
  mutate(n_adj = nlcd_downsample*n) %>%
  inner_join(nlcd_types, by = 'code') %>%
  mutate(sqmi = n_adj * 0.00347492) %>% # 30m^2 -> sq mile
  ggplot() +
  geom_col(aes(fill=description, x=category, y=sqmi), position='stack') +
  scale_fill_manual(values = nlcd_colors) + 
  geom_text(aes(x=category, group=code, y=sqmi, label=description), size=3, 
          position = position_stack(vjust = 0.5)) +
  scale_y_continuous(label=comma) +
  ylab("Sq miles") + 
  xlab("NLCD Category") +
  ggtitle('NLCD land type within 2 miles of NY substations / 1 mile of 230kV+ lines', 
          subtitle = "Win Climate, 2022") +
  watermark("Draft") +
  theme(legend.position="none")
```

Not restricted to buffer around stations

```{r}
nlcd %>%
  pull(nlcd2016_ny.tif) %>%
  st_as_stars(downsample = nlcd_downsample) %>%
  table() %>%
  as_tibble() %>%
  rename(code = 1) %>%
  mutate(n_adj = nlcd_downsample*n) %>%
  inner_join(nlcd_types, by = 'code') %>%
  mutate(sqmi = n_adj * 0.00347492) %>% # 30m^2 -> sq mile
  ggplot() +
  geom_col(aes(fill=description, x=category, y=sqmi), position='stack') +
  scale_fill_manual(values = nlcd_colors) + 
  geom_text(aes(x=category, group=code, y=sqmi, label=description), size=3, 
          position = position_stack(vjust = 0.5)) +
  scale_y_continuous(label=comma) +
  ylab("Sq miles") + 
  xlab("NLCD Category") +
  ggtitle('NLCD land type across NY State', 
          subtitle = "Win Climate, 2022") +
  watermark("Draft") +
  theme(legend.position="none")
```

```{r, fig.width=10}
landuse_buffer <- nlcd[buffer_together] %>%
  pull(nlcd2016_ny.tif) %>%
  st_as_stars(downsample = nlcd_downsample) %>%
  table() %>%
  as_tibble() %>%
  rename(code = 1) %>%
  mutate(n_adj = nlcd_downsample*n) %>%
  inner_join(nlcd_types, by = 'code') %>%
  mutate(sqmi_buffer = n_adj * 0.00347492) # 30m^2 -> sq mile

landuse <- nlcd %>%
  pull(nlcd2016_ny.tif) %>%
  st_as_stars(downsample = nlcd_downsample) %>%
  table() %>%
  as_tibble() %>%
  rename(code = 1) %>%
  mutate(n_adj = nlcd_downsample*n) %>%
  inner_join(nlcd_types, by = 'code') %>%
  mutate(sqmi = n_adj * 0.00347492) # 30m^2 -> sq mile

landuse %>%
  inner_join(landuse_buffer, by = 'code') %>%
  mutate(description = description.x, 
         category = category.x) %>%
  filter(!category %in% c("Water", "Wetlands", "Developed")) %>%
  mutate(total_sqmi = sum(sqmi)) %>%
  group_by(category) %>%
  summarize(buffer_sqmi = sum(sqmi_buffer), 
            category_sqmi = sum(sqmi),
            total_sqmi = first(total_sqmi),
            pct_category_in_buffer = buffer_sqmi / category_sqmi,
            pct_in_buffer_nys = buffer_sqmi / total_sqmi,
            pct_300sqmi = 300/buffer_sqmi)

```
If we don’t want to cut down trees, only 10% of undeveloped NYS land is buildable (i.e. rural land close to grid). 

That represents a third of all the agricultural land.

## Slope

```{r}
library(raster)
```